---
title: "Inverse DO model"
author: "Alice Carter"
date: "6/16/2020"
output: html_document
---
Version 1

Estimation of dissolved oxygen through inverse modeling
Based on Notes from Gaby Katul

**Basic Equation:**  
Consider an extensive stream where the flow is uniform.  The mass balance equation for oxygen is
$$\frac{\delta VC}{\delta t} = -(\frac{\delta q_{x}}{\delta x}+\frac{\delta q_{y}}{\delta y}+\frac{\delta q_{z}}{\delta z})+S$$
where t is time, V is the control volume, C is the pointwise oxygen concentration, q_x, q_y,and q_z are the mass fluxes in the x (longitudinal), y (lateral), and z (vertical directions) respectively (due to advective and dispersive transport), and S are all the local sources and sinks for oxygen (biotic and abiotic).  

**Assumptions:**   
Uniform flow implies that   
$$\frac{\delta q_{x}}{\delta x}+\frac{\delta q_{y}}{\delta y}=0$$

Integrate with respect to z and define fluxes per unit ground area A assuming the channel is rectangular of depth H:  
$$\frac{\delta H\overline{C}}{\delta t} = q(H)-q(0)+ \int^{H}_{0} Sdz $$
where $q(H)$ is the flux of oxygen with the atmosphere, $\overline{C}$ is the depth-averaged concentration, and the term  $B(t)=-q(0)+\int^{h}_{0} Sdz$ represents all production and removal of oxygen (say by plants, mircobes, fish, sediments, etc…).  

**The model**  
We seek a describtion of $B(t)$ from $\overline{C}(t)$ using
$$\frac{\delta H\overline{C}}{\delta t} = q(H) + B(t).$$

To do so, a link between $q(H)$ and $C$ is required. One model is a bulk gas transfer given by
$$q(H)=k(C*-\overline{C}),$$
where $C*$ is the oxygen saturation (varies with water temperature), and $k$ is the gas transfer velocity that can be inferred from a number of turbulence theories. If so, then
$$\frac{\delta H\overline{C}}{\delta t} = k(C*-\overline{C})+B(t).$$

Provided that water depth, concentration, $k$, and water temperature are available, $B(t)$ can be inferred and analyzed for different environmental conditions (e.g. as a function of light, temperature, or other chemical compounds such as N from fertilizer runoff, etc….)  
The micro-eddy model (in the absence of bubbles) stipulates that $k = c1(Sc)^{-1/2}(\nu \epsilon)^{1/4},$ where $c1$ is a coefficient, $Sc$ is molecular Schmidt number for oxygen, $\nu$ is the kinematic viscosity of water, and $\epsilon$ is the turbulent kinetic energy dissipation rate.  

In the absence of wind at the water surface, the $\epsilon = c2(u_{*})^{3}/H$, where $c2$ is a similarity coefficient and $u_{*}$ is the friction velocity.  
Last, the friction velocity can be determined from $u_{*}=\sqrt{gHS_{0}}$, where $g$ is the gravitational acceleration and $S_{0}$ is the bed slope.


**Revisions:**  
Revisions can be made to include large eddies, bubbles, and winds – but my preference is to start simple.  

So, with c1 and c2 determined from laboratory studies, $B(t)$ can be determined for all rivers.  
It is emphasized that the $B(t)$ is only an overall bulk, aggregate, or effective source/sink term and must be viewed as a lumped quantity.  However, if the goal is to distinguish why different ecosystems  experience differing patterns of oxygen fluctuations, it may be sufficient – as it links measured concentration time series, water level, water temperature, and channel properties (e.g. $S_{0}$) to overall sources or sinks (or production/consumption of oxygen). 

The final summary result:
$$B(t)=\frac{d(H\overline{C})}{dt}-c1(Sc)^{-1/2}(c2\nu H^{1/2}(g S_{0})^{3/2})^{1/4}(C^{*}-\overline{C}) $$

```{r setup, include=FALSE}
library(readr)
library(fitdistrplus)
library(lubridate)

setwd(hypox_projdir)

#######################################
#read in data and metadata
sites <- read_csv("NHC_map/NC_synopticSamplingSites.csv")
annual_dat <- read_csv("data/raw/2019SPsites.csv")
annual_dat$DateTime <- with_tz(annual_dat$DateTime_UTC, tz="EST")
annual_dat$date <- as.Date(annual_dat$DateTime, tz = "EST")
annual_dat$year <- year(annual_dat$date)
annual_dat<- annual_dat[annual_dat$year %in% 2017:2019,]
annual_dat$site_year <- paste(annual_dat$site,year(annual_dat$date), sep="_")

dat <- annual_dat[annual_dat$site_year=="NHC_2019",]
```



Test model on one year of data from New Hope Creek, this can be compared to the model output from Stream Metabolizer

First initialize constants
$\nu$ is the kinematic viscosity of water, which equals $\eta /\rho$, or the dynamic viscosity divided by the density of water 

$Sc$ is the schmidt number for $O_{2}$, which 
```{r include=FALSE}

