---
title: "Watershed Delineation"
author: "Mark Hagemannr"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r}
# update and load packages
# library(devtools)
# install_github('markwh/streamstats')
library(streamstats)
library(dplyr)
```

Assuming you already have the lat-lon of an outlet, computing the contributing watershed is as easy as:

```{r}
sites <- read.csv("NC_synopticSamplingSites.csv", header=T)
MCout <- sites[sites$site=="MCconf",2:3]
NHCout <- sites[sites$site=="Blands",2:3]

MC <- delineateWatershed(xlocation = MCout$Long, ylocation = MCout$Lat, crs = 4326, 
                          includeparameters = "true", includeflowtypes = "true")
NHC <- delineateWatershed(xlocation = NHCout$Long, ylocation = NHCout$Lat, crs = 4326, 
                          includeparameters = "true", includeflowtypes = "true")

```

(if timeout is reached before a result is returned, you can specify it using `setTimeout([number of seconds])`)

Here, `crs` is the coordinate reference system number (ESPSG spatial reference code). 

To see what it returned, the `leafletWatershed` function gives a simple interactive map.

```{r}
leafletWatershed(MC)
leafletWatershed(NHC)
```

Other statistics can be found using the `computeChars` (for watershed characteristics such as basin area and land-use) and `computeFlowStats` (for statistics such as flow percentiles)

```{r}
MCchars <- computeChars(workspaceID = MC$workspaceID, rcode = "NC")
MCchars$parameters

NHCchars <- computeChars(workspaceID = NHC$workspaceID, rcode = "NC")
NHCchars$parameters

WatershedChars <- MCchars$parameters %>% rename(MC=value)
WatershedChars <- full_join(WatershedChars, NHCchars$parameters[,c(4,6)], by = "code") %>%
  rename(NHC=value) %>% select(-ID)

#Compute drainage area in km2
xx <- WatershedChars[WatershedChars$code=="DRNAREA", c(5,6)]
area <- data.frame(c("Drainage Area km2","Drainage area in km2", "", "square kilometers",  xx*(1.60934^2)), stringsAsFactors = F) 
names(area) <- (colnames(WatershedChars))
# Slope in m/km and percent
xx <- WatershedChars[WatershedChars$code=="CSL10_85fm", c(5,6)]
slopeper <- data.frame(c("Channel Slope Percent","10 to 85 channel slope in percent", "", "percent", xx*(100/5280)),stringsAsFactors = F)
names(slopeper) <- (colnames(WatershedChars))
slopemkm<- data.frame(c("Channel Slope m/km","10 to 85 channel slope in m per km", "", "m per km", xx*(12*2.54/100/1.60934)),stringsAsFactors = F)
names(slopemkm) <- (colnames(WatershedChars))


binddf = rbind(area,slopeper,slopemkm )
WatershedChars <- rbind(WatershedChars, binddf)

write.csv(WatershedChars, file = "NHCMCwatershed_characteristics.csv", row.names = FALSE)
```

```{r}
NHCstats <- computeFlowStats(workspaceID = NHC$workspaceID, rcode = "NC")
NHCstats$BANKFULL$flow

MCstats <- computeFlowStats(workspaceID = MC$workspaceID, rcode = "NC")
MCstats$BANKFULL$flow
```


